---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
movies <- read_csv("data/movies.csv")
```

# How could we clean this data?

We could assign release_date to a numeric value for easier analysis?

```{r}
glimpse(movies)
head(movies)
```

# 2. Select specific columns

```{r}
movies_selected <-
  movies %>% 
  select(title, runtime, budget)
movies_selected
```

# 3. Count the number of missing values
Ok, now take your subsetted data movies_selected, and count the number of missing values in each column.

```{r}
movies_selected %>% 
  filter(is.na(title))
```

```{r}
movies_selected %>% 
  filter(is.na(runtime))
```

```{r}
movies_selected %>% 
  filter(is.na(budget))
```

There’s not too many NAs, but there are actually quite a few movies with a runtime of 0 in this dataset. Since it’s impossible for a movie to have a runtime of 0, use the function na_if to convert runtimes of length 0 into NAs. Save this into a new variable called movies_runtime. Then count how many missing values there are now in the column runtime.

# 4. Are there hidden NAs?

```{r}
movies_selected %>% 
  filter(runtime  <= 1)
```


```{r}
movies_runtime <-
  movies_selected %>% 
  mutate(runtime = na_if(runtime, 0)) %>% 
  arrange(runtime)
```

```{r}
movies_runtime %>% 
  filter(is.na(runtime))
```


# 5. Impute Missing Values

Impute the missing values in runtime with the median runtime using coalesce and save this into a new variable called movies_imputed. There should now be no missing values in the column runtime - check this to make sure.

```{r}
movies_imputed <- 
  movies_runtime %>% 
  mutate(runtime = coalesce(runtime, median(runtime, na.rm = TRUE)))
```

```{r}
movies_imputed %>% 
  count(is.na(runtime))

movies_imputed %>% 
  filter(is.na(runtime))
```

# 6. Slice

Finally, use the slice_min and slice_max functions on movies_imputed to get the movies with the 10 shortest and 10 longest runtimes.

```{r}
movies_imputed %>% 
  slice_min(runtime, n=10)
```

```{r}
movies_imputed %>% 
  slice_max(runtime, n=10)
```

# 7. Using ifElse

Using the if_else function, impute anything with a budget below $100 with the median budget. Overwrite your movies_imputed so it contains the new budget values.

```{r}
movies_imputed %>% 
  mutate(budget = if_else(budget < 100, median(budget), budget))
```

# Extension 1
Using case_when, create a new column called budget_type and label budgets as follows:

Budgets below 12 million as ‘Small budget’
Budgets between 12 and 40 million as ‘Medium budget’
Budgets 40 millions and above as ‘Big budget’
Save this into a new variable called movie_budgets.

```{r}
movie_budgets <- 
  movies_imputed %>% 
  mutate(budget_type =
           case_when(
             budget < 12e6 ~ "Small budget",
             budget < 40e6 ~ "Medium budget",
             TRUE ~ "Big budget"
           )) %>% 
  arrange(budget)
```

# Extension 2

Using across and where, summarise the number of missing values, first across all columns of type character, and then across all columns of type numeric.

```{r}
movies %>% 
  filter(
    is.na(original_title))

?across
```

```{r}
movies %>% 
  filter(is.na
         (across(
           where(is.character)
         )))
```

```{r}
movies %>% 
  filter(is.na
         (across(
           where(is.numeric)
         )))
```


