---
title: "R Notebook"
output: html_notebook
---

Distributions homework
You are given a cut of a database containing details of mobile phones and their customer reviews on an online shopping site in file 20190928-items.csv. Each row in the data corresponds to one model of mobile phone. Use the data to answer the following questions.


As part of the MVP we want you not to just run the code but also have a go at intepreting the results - this is the main learning objective from this homework. Don’t worry if you aren’t sure, it’s ok to get it wrong, but give it a go, through trying we will learn.

1 MVP
1.1 Question 1
Load the data, clean_names() and explore it.

```{r}
library(tidyverse)
library(janitor)

phones <- read_csv("data/20190928-items.csv") %>% 
  # convert prices to numeric
  mutate(prices = as.numeric(str_replace_all(prices, "\\$", ""))) %>% 
  clean_names() %>% 
  select(asin, title, brand, rating, total_reviews, prices)
```

```{r}
# What are the minimum and maximum values for number of reviews?
range(phones$total_reviews)
```

We have phones rated from once to 984 times. We could say that the reviews are more meaningful if they are from a larger number of reviewers.

```{r}
# How many phones have less than 5 ratings?
phones %>% 
  filter(total_reviews < 5)
```
139 of our 792 observations have less than five ratings.

```{r}
# How many phones have less than 2 ratings?
phones %>% 
  filter(total_reviews < 2)
```
63 have less than 2 ratings.

```{r}
# are there any repeat models in the data?
length(unique(phones$asin))
```
We have 792 unique asins, which look like a model number, so I don't think we have any repeated observations.

```{r}
# check this with phone title
length(unique(phones$title))
```

We only have 785 unique titles for the phones suggesting 7 overlaps

```{r}
phones[duplicated(phones$title),]
```

# Clean instances of these titles, merging the rows by edditing rating as average of the two values and total_reviews as the sum and keeping the asin of the most reviewed one

# Also clean titles using regex to put all models of one phone together, not easy in this dataset.


1.2 Question 2
Find the brand with the highest numbers of phones in the dataset.
Samsung has the greatest number of phones in this dataset. 397, compared to 101 by the second greatest, Apple.

```{r}
phones %>% 
  group_by(brand) %>% 
  summarise(num_phones = n()) %>% 
  ggplot() +
  aes(x = reorder(brand, -num_phones), y = num_phones) +
  geom_col()
```

```{r}
top_brand <- phones %>% 
  group_by(brand) %>% 
  summarise(num_phones = n()) %>%
  arrange(desc(num_phones)) %>% 
  head(1) %>% 
  pull(brand)
```

1.3 Question 3
For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

```{r}
top_brand_phone <- phones %>% 
  filter(brand == top_brand)

top_brand_phone_stats <- top_brand_phone %>% 
  summarise(n = n(),
            mean = mean(rating),
            sd = sd(rating))

top_brand_phone %>% 
 # mutate(binned = cut(rating, breaks = 25)) %>% 
  filter(total_reviews > 2) %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
#  geom_smooth(aes(y = ..density..), col = "dark red") +
  stat_function(
    fun = dnorm, 
    args = list(
      mean = top_brand_phone_stats$mean, 
      sd = top_brand_phone_stats$sd
    ),
    col = "steel blue"
  )
```
Our data fits fairly well to a normal distribution curve, skewe left and with outliers at ratings of 1 and 5, which can be removed by filtering for phones which have less than 2 reviews (where a rating of 1 or 5 is possible based on only one person's opinion).
Ratings peak around 3.7 whereas our mean is around 3.5.



1.4 Question 4
We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

