---
title: "R Notebook"
output: html_notebook
---

# Starbucks Calorie Dataset Exercise

Data on Calories in Starbucks drinks from kaggle

# Necessary libraries

```{r}
library(tidyverse)
```

# 1. Import and get to know the Data

```{r}
starbucks_calories <- read_csv("data/starbucks_drinkMenu_expanded.csv")

dim(starbucks_calories)

names(starbucks_calories)

glimpse(starbucks_calories)

class(starbucks_calories)
```

Starbucks Calories is a dataframe containing 242 observations or rows and 18 variables.
Starbucks Calories has variables containing information about each drink, its category and preparation style and its nutritional content including calories, sugar, fats, etc.
Our percentage values have imported as character vectors as they have the percentage symbol in the field. The other numeric values are numbers and the descriptive fields are character vectors


# 2. Investigating Calories

Reduce data to look at specific columns using select, then check for NAs in the Calories column (none found)



```{r}
select(starbucks_calories, Beverage_category, Beverage, Beverage_prep, Calories)

filter(starbucks_calories, is.na(Calories))
```

# 3. Find out which drinks have more calories than 10 cups of sugar

10 cups of sugar = 135 calories, so we are exploring whether each drink has more or less than 135 calories. We'll store this in a column called calorie_diff. We'll say if each drink has more or less calories than 10 cups of sugar in a column called more_sugar_than_ten_cups

```{r}
starbucks_calories %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>%
  mutate(calorie_diff = 135 - Calories) %>% 
  mutate(more_sugar_than_ten_cups = calorie_diff > 0)
```


# 4. Exploring mean calories by beverage_category

We will now group the drinks by the Beverage Category provided in the dataset and find the mean for each group.

```{r}
starbucks_calories %>% 
  names()
?summarise

```

```{r}
starbucks_calories %>% 
  group_by(Beverage_category) %>% 
  summarise(mean_calories = mean(Calories), n=n())
```

# Highest and Lowest Calories

```{r}
calories_ordered <- starbucks_calories %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>%
  arrange(Calories)

  head(starbucks_calories, 3)
```


```{r}
tail (starbucks_calories, 3)
```

# 5. Average number of calories in each Beverage Prep Category

```{r}
starbucks_calories %>% 
  group_by(Beverage_prep) %>% 
  summarise(mean_calories = mean(Calories), n=n())
```

# 6. Which Beverage Prep contains more than the average calories of all drinks?

Find the mean for all drinks using summarise.

```{r}
mean_calories <-
  starbucks_calories %>% 
  summarise(calories = mean(Calories))
```

Mean of all drinks = 193.871, now stored as the variable mean_calories.

Then create a new column called more_than_average which returns TRUE if Calories 

```{r}
calories_ordered %>% 
  group_by(Beverage_prep) %>% 
#   mutate(more_than_average = (Calories - mean_calories) <= 0)
  mutate(more_than_average = Calories > pull(mean_calories))  
```


# 7. Which is the best type of coffee drink to get if you’re worried about consuming too many calories?

Find out which beverage types have the lowest average calories

```{r}
starbucks_calories %>% 
  group_by(Beverage_prep) %>% 
  summarise(mean_calories = mean(Calories), n=n()) %>% 
  arrange(mean_calories) %>% 
  head(10)
```

If you are worried about your calory consumption, you are best off drinking from the Solo, Doppio and Short range because these are the 3 beverage types with the lowest average calories.
These are all categories which describe the size of drink so this implies that the smaller your drink the less calories you will be taking in.
When we look at the 10 beverage types with the lowest we see that nonfat milk appears 3 times, so we can also recommend that you use nonfat milk in whatever drink you choose.



