---
title: "R Notebook"
output: html_notebook
---

# Starbucks Calorie Dataset Exercise

Data on Calories in Starbucks drinks from kaggle

# Necessary libraries

```{r}
library(tidyverse)
```

# 1. Import and get to know the Data

```{r}
starbucks_calories <- read_csv("data/starbucks_drinkMenu_expanded.csv")

dim(starbucks_calories)

names(starbucks_calories)

glimpse(starbucks_calories)

class(starbucks_calories)
```

Starbucks Calories is a dataframe containing 242 observations or rows and 18 variables.
Starbucks Calories has variables containing information about each drink, its category and preparation style and its nutritional content including calories, sugar, fats, etc.
Our percentage values have imported as character vectors as they have the percentage symbol in the field. The other numeric values are numbers and the descriptive fields are character vectors


# 2. Investigating Calories

Reduce data to look at specific columns using select, then check for NAs in the Calories column (none found)



```{r}
select(starbucks_calories, Beverage_category, Beverage, Beverage_prep, Calories)

filter(starbucks_calories, is.na(Calories))
```

# 3. Find out which drinks have more calories than 10 cups of sugar

10 cups of sugar = 135 calories, so we are exploring whether each drink has more or less than 135 calories. We'll store this in a column called calorie_diff. We'll say if each drink has more or less calories than 10 cups of sugar in a column called more_sugar_than_ten_cups

```{r}
starbucks_calories %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>%
  mutate(calorie_diff = 135 - Calories) %>% 
  mutate(more_sugar_than_ten_cups = calorie_diff > 0)
```


# 4. Exploring mean calories by beverage_category

We will now group the drinks by the Beverage Category provided in the dataset and find the mean for each group.

```{r}
starbucks_calories %>% 
  names()
?summarise

```

```{r}
starbucks_calories %>% 
  group_by(Beverage_category) %>% 
  # summarise, need to use mean function, but not enough time to figure out
  summarise(n=n())
```

# Highest and Lowest Calories

```{r}
calories_ordered <- starbucks_calories %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>%
  arrange(Calories)

  head(starbucks_calories, 3)
```


```{r}
tail (starbucks_calories, 3)
```


5 Question 5
Let’s look at this a different way. What is the average number of calories in each Beverage_prep type?


6 Question 6
Which Beverage_prep type contains more than the average calories of all drinks?
Hint: to answer this, you’ll have to first figure out what the average calories across all drinks are, and then use that as a filter for the grouped Beverage_prep data.

7 Question 7
Which is the best type of coffee drink to get if you’re worried about consuming too many calories?