---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```


```{r}
new_coders <- read_csv("data/new_coders.csv")
```

# Sampling
Taking a random set of data

Sample_n - Sample by number

```{r}
new_coders %>% 
  sample_n(5)
```

Sample_frac - Sample by proportion

```{r}
new_coders %>% 
  sample_frac(0.01)
```

# Distinct

Keep just the ones you called

```{r}
new_coders %>% 
  distinct (age, gender)
```

Keep all values

```{r}
new_coders %>% 
  distinct (age, gender, .keep_all = TRUE)
```

# Slicing

Subsetting by position not value

First 5

```{r}
new_coders %>% 
  slice(1:5)
```

2nd, 3rd, 5th and 10th

```{r}
new_coders %>% 
  slice(c(2,3,5,10))
```

Slice using group by will give you a slice from each of the variables we group by

```{r}
new_coders %>% 
  group_by(job_pref) %>% 
  slice (1:5)
```

# Slice min and max

Select the minimum and maximum rows for the columns selected

```{r}
new_coders %>% 
  slice_max(last_yr_income, n =5)
```

```{r}
new_coders %>% 
  slice_min(last_yr_income, n = 5)
```

But this brings the all the rows containing the 5 lowest values.

```{r}
new_coders %>% 
  slice_min(last_yr_income, n = 5, with_ties = FALSE)
```

We can use slice min and max with group by

```{r}
new_coders %>% 
  group_by(job_pref) %>% 
  slice_max(expected_earn, n=1, with_ties = FALSE)
```

