---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
tweets <- read_csv("data/code_clan_tweets.csv")

names(tweets)

dim(tweets)
```


Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
tweets %>% 
  filter(is_quote = TRUE) %>% 
  summarise(number_of_favourited_tweets = sum(favorite_count))
```


Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
tweets %>% 
  filter(is_quote = TRUE) %>% 
  group_by(source) %>% 
  summarise(mean_by_source = mean(retweet_count))
```


Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

I did this and then changed it to mean as this is a more useful figure given that we have different numbers of media types.

```{r}
# Assume that any tweets without a listed media type are of type “text”.

tweets_media <- tweets %>%
  mutate(media_type = if_else(media_type == "", "text", media_type),
         media_type = if_else(is.na(media_type), "text", media_type)
  ) %>% 

# split by media type
  group_by(media_type) %>% 
  
# Count the total number of likes (i.e. total of favorite_count)
  summarise(num_likes = mean(favorite_count, na.rm = FALSE)) %>% 

# arrange them from most likes to least.
  arrange(desc(num_likes))

tweets_media
  
```


Find the mean number of characters that a CodeClan tweet contains.
234

```{r}
tweets %>% 
  select(text) %>% 
  mutate(tweet_length = length(text)) %>% 
  summarise (avg_tweet = mean(tweet_length))
```
2#\

The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
status <- read_csv("data/code_clan_info.csv")
```

```{r}
tweet_status <- left_join(
  tweets, status,
  by = "tweet_id"
)
```


From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags.

```{r}
codeclan_hashtags <- tweet_status %>%
  select(tweet_id, hashtags) %>% 
  mutate(hashtags = str_to_lower(hashtags)) %>% 
  filter(!is.na(hashtags))
```


4 Extensions
Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.
Hint

```{r}
codeclan_hashtags %>% 
  filter(str_detect(hashtags, "c\\("))
```


Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.
Hint
```{r}
codeclan_hashtags %>% 
  str_detect("[Ed]dinburgh")
```


Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.
Hint
