library(tidyverse)
library(janitor)

```{r}

oj_data <- read_csv("data/orange_juice.csv") %>% 
  clean_names()

```


# Exploration and Understanding

Purchases of two brands of orange juice - target variable - and data on week of purchase, store, price of CH brand, price of MM brand, discount applied to either brand, whether each brand was on sale (equal to is there a value in discount), brand loyalty for CH (and not MM), sale price of each brand (equal to price less discount), difference in price (equal to sale price - sale price), store 7 factor (not helpful), difference in list price (equal to price_mm - price_ch), percentage discount for each brand (alias), store (alias storeid)


# Remove aliases

Note that removing speical will lose any indication of whether a promotional price affects sales, only retaining how the percentage discount affects sales.
*Consider re-rerunning with special_ch and special_mm in place of pct_disc_ch and pct_disc_mm*

```{r}
oj_reduced_var_data <- oj_data %>% 
  select(-c(
  # Remove both price_, which is captured in sale price and percentage discount
    price_ch, price_mm, 
  # Remove both disc_, discount amounts, which is captured in percentage discount and sale price
    disc_ch, disc_mm,
  # Remove special, which is captured in percentage discount 
    special_ch, special_mm,
  # Remove price difference (derived from sale_price_)
    price_diff,
  # Remove store7 factor and store (contained in storeid)
    store7, store
  ))
```


# Convert variables to factors

Time of purchase may be a predictor variable, for example if the brand has run advertising, or has had negative press coverage. The Data contains 51 weeks worth of observations, so I am choosing to convert this to a series of 12 periods which will give us a rough idea of month of purchase, although we will not know which months we are looking at and the seperations will not be equal.
This will still give us an indication of whether date of purchase is a useful predictor and can inform future data collection.

```{r}
oj_reduced_var_data_factored <- oj_reduced_var_data %>% 
  mutate(
    purchase_ch = as_factor(if_else(purchase == "CH", 1, 0)),
    month_of_purchase = cut(weekof_purchase, breaks = 12, labels = 1:12),
    loyal_ch = as.numeric(format(round(loyal_ch, 1), nsmall = 1))
  ) %>% 
  select(-c(purchase, weekof_purchase)) %>% 
  mutate(across(where(is_numeric), as_factor))
```

# Check for aliases
```{r}
# don't run, lot's of printout.
# alias(purchase_ch ~ ., data = oj_reduced_var_data_factored)
```

# Write clean data
```{r}
write_csv(oj_reduced_var_data_factored, "../data/clean_oj.csv")
```

