
```{r}
library(caret)
library(glmulti)
library(tidyverse)
```
# Load in clean data and refactor the numeric variables
```{r}
oj_data <- read_csv("../data/clean_oj.csv") %>% 
  mutate(across(where(is_numeric), as_factor))
```


# Create train and test datasets.
```{r}
train_index <- createDataPartition(
  oj_data$purchase_ch, 
  p = 0.85, 
  list = FALSE, 
  times = 1)

training_data <- oj_data[train_index, ]
test_data <- oj_data[-train_index, ]
```

Check purchases in training and test set are comparable
   Both sets have similar distribution of purchases.
```{r}
ggplot(training_data) +
  aes(x = purchase_ch) +
  geom_bar()

ggplot(test_data) +
  aes(x = purchase_ch) +
  geom_bar()
```

# Predictor selection for model

Check size of request to glmulti 

```{r}
glmulit_fit <- glmulti(
  # include all variables
  purchase_ch ~ .,
  # include interactions between variables
  data = training_data,
  level = 2,
  maxsize = 10,
  marginality = TRUE,
  # set trial run
  method = "d",
  confsetsize = 500,
  plotty = FALSE,
  report = TRUE,
  fitfunction = "glm",
  family = binomial(link = "logit"),
  # measure models on AIC as requested in brief
  crit = "aic"
)

summary(glmulit_fit)
```
107972 possible models. Should be possible in about 10-15 minutes.

```{r}
glmulit_fit <- glmulti(
  # include all variables
  purchase_ch ~ .,
  # include interactions between variables
  data = training_data,
  level = 2,
  maxsize = 10,
  marginality = TRUE,
  # use algorithm method
  method = "g",
  confsetsize = 10,
  plotty = FALSE,
  report = TRUE,
  fitfunction = "glm",
  family = binomial(link = "logit"),
  # measure models on AIC as requested in brief
  crit = "aic"
)

summary(glmulit_fit)
```

After 90 generations:
Best model: purchase_ch~1+store_id+loyal_ch+pct_disc_ch+list_price_diff
Crit= 731.641583088339
Mean crit= 739.62794186827
Change in best IC: 0 / Change in mean IC: -1.09340887670089

# Build best model
```{r}
model <- glm(purchase_ch ~ , 
             data = training_data, 
             method = "lm", family = binomial(link = "logit"))
```


# Test model on test_data

```{r}
predictions <- predict(model, test_data)

predictions - test_data$purchase_ch
```
