---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(CodeClanData)

backpacks <- backpack
```

In all plots you make, ensure your labels are appropriate. You also must write down a short summary of what each plot tells you in order to meet MVP.

# MVP


Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

Contains backpack weight, body weight, what I assume to be a ratio between these two, what looks like a logical column for back problems, the person's major, their year (at university/school?), sex, status (which might refer to Undergraduate and Graduate) and a variable called units which could be how many subjects they are studying.

I assume the weights are given in pounds.

Units range from 0-19.
Year from 0-6.
Majors contains 41 values which are organised as factors.

```{r}
head(backpacks)

unique(backpacks$major)
```


Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

This line graph shows theferquency of the occurence of a weight of backpack in the dataset. It shows that backpack weight has a standard distribution centring just under 10 pounds and tailing on the heavy side. 

```{r}
ggplot(backpacks) +
  aes(x = backpack_weight) +
  geom_density() +
  labs(
    title = "Frequency of backpack weights carried by higher education students",
    y = "Frequency",
    x = "Backpack Weight (pounds)"
  )
```

Question 3.

Are there differences in backpack weight between sex?

This graph shows the frequency of different backpack weights split by the student's gender.
The backpack weight for women trends slightly smaller than it does for men centring just below 10 pounds and tailing on the heavy side.
The backpack weight for men trends slightly larger than womens and centres at 10 pounds with a longer tail to the heavier side.

```{r}
ggplot(backpacks) +
  aes(x = backpack_weight, colour = sex) +
  geom_density() +
  labs(
    title = "Frequency of backpack weights for each gender",
    y = "Frequency",
    x = "Backpack Weight (pounds)"
  )
```

Question 4.

What is the relationship between body weight and backpack weight?

```{r}
ggplot(backpacks) +
  aes(x = backpack_weight, y = body_weight) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(backpacks) +
  aes(x = backpack_weight, y = body_weight) +
  geom_point() +
  geom_smooth()
```


Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

Question 6.

What proportion of students report back pain?

Question 7.

Is this proportion different between men and women?

Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

3 Extensions


Question 1.

Does reported backpain differ across the year groups? Ignore year 0 in your data.

Question 2.

Is there a difference between how heavy the bags those in different years carry?