---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
revenue <- qb_revenue_breakdown
monthly <- qb_monthly_sales
competitors <- qb_competitors
device <- qb_device_data
```

4.1 Part 1
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
revenue %>% 
  ggplot() +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col()
```


Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
monthly %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = Date, y = Amount) +
  geom_line()
```


Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.
```{r}
monthly %>% 
  ggplot() +
  aes(x = Date, y = Amount, group = Cashflow, colour = Cashflow) +
  geom_line()
```


Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.
```{r}
competitors %>% 
  ggplot() +
  aes(x = Year, y = Revenue, group = Company, colour = Company, fill = Company) +
  geom_area()
```


Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
competitors %>% 
  ggplot() +
  aes(x = Year, y = Revenue, group = Company, colour = Company, fill = Company) +
  geom_line() +
  geom_point()
```


Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}

# One plot per person showing the 4 devices, with qikbit as a colour and the rest as grey, shows that qikbit tracks similarly to its competitors.

device %>% 
  mutate(observation = row_number()) %>% 
  mutate(time = 60 * hours + mins) %>% 
  ggplot() +
  aes(x = observation, y = time, group = device) +
  geom_line(stat = "identity") +
  facet_grid(~ device)

device %>% 
  mutate(observation = row_number()) %>% 
  mutate(time = 60 * hours + mins) %>% 
  ggplot() +
  aes(x = time, y = counts) +
  geom_line() +
  facet_grid(id ~ device)
```

